### 2022-12-31 references/yellowpaper/JS.tex
Add dummy 1.6 method

\section{Javascript API}\label{app:jsapi}

The JavaScript API provides a consistent API across multiple scenarios including each of the clients' web-based in-process \DH{}App frameworks and the out-of-process RPC-based infrastructure. All key access takes places though the special \texttt{eth} object, part of the global namespace.

\subsection{Values}
There are no special object types in the API; all values are strings. As strings, values may be of several forms, and are interpreted by the API according to a series of rules:

\begin{enumerate}
\item If the string contains only digits from 0-9, then it is interpreted as a decimal integer;
\item if the string begins with the characters \texttt{0x}, th